---
title: "Kinetic analysis of yeast-yeast interactions in oenological conditions"
author: "thibault nidelet"
date: "05/06/2019"
output: html_document
---

# README

Here you can find the script for all analysis and graphiqual representation of the article "Kinetic analysis of yeast-yeast interactions in oenological conditions"

We used two dataset for our analysis :  

* CO2.csv  
* all_data.csv  

The first dataset (**CO2.csv**) contain the evolution of CO2 production every 20mn up to 300h for each fermentation.
The columns are :  

* Blocs (the 5 different blocks of the experiment)  
* Fermenter (the number of the fermenter of each fermentation. The same fermenters can be found in different blocks)  
* Species (the species composition at t0 of each fermentation)  
* Time (the time in hours)  
* CO2 (the production of CO2 in g/L)  
* CO2_rate (the rate of CO2 production in g/L/h)  


The second dataset gives population and metabolites concentration for around 10 points across the fermentation. The columns are : 

* Blocs (the 5 different blocks of the experiment)  
* Fermenter (the number of the fermenter of each fermentation. The same fermenters can be found in different blocks)  
* Species (the species composition at t0 of each fermentation)  
* Time (the time in hours)  
* CO2 (the production of CO2 in g/L)  
* Pop_total (the total number of cells, alive or dead measure by cytometer in cells/mL)  
* Prop_Sc (the proportion of the species S. cerevisiae, between 0 and 1)  
* viability (the viability between 0 and 1)  
* glucose (the concentration of glucose in g/L measured by HPLC)  
* fructose (the concentration of fructose in g/L measured by HPLC)  
* glycerol (the concentration of glycerol in g/L measured by HPLC)  
* acetate (the concentration of acetate in g/L measured by HPLC)  
* ethanol (the concentration of ethanol in g/L measured by HPLC)  
* alpha (the concentration of alpha ketoglutarate in g/L measured by HPLC)  
* succinate (the concentration of succinate in g/L measured by HPLC)  
* pyruvate (the concentration of pyruvate in g/L measured by HPLC)  
* Ammonium (the concentration of ammonium in mg/L measured by HPLC)  
* Alanine (the concentration of alanine in mg/L measured by HPLC)  
* Arginine (the concentration of arginine in mg/L measured by HPLC)  
* Aspartate (the concentration of aspartate in mg/L measured by HPLC)  
* Glutamine (the concentration of glutamine in mg/L measured by HPLC)  
* Glutamate (the concentration of glutamate in mg/L measured by HPLC)  
* Glycine (the concentration of glycine in mg/L measured by HPLC)  
* Histidine (the concentration of histidine in mg/L measured by HPLC)  
* Methionine (the concentration of methionine in mg/L measured by HPLC)  
* Phenylalanine (the concentration of phenylalanine in mg/L measured by HPLC)  
* Serine (the concentration of serine in mg/L measured by HPLC)  




# Loading packages and Datasets
```{r }
library(tidyverse)
library(ggthemes)
library(growthcurver)
library(stringr)
library(patchwork)
library(Rmisc)
library(agricolae)
library(ggpubr)
library(officer)
library(rvg)


erreur.std <- function(x)
{ x <- na.omit(x)
ET <- sd(x)/sqrt(length(x))
return(ET)
}

#loading the data of CO2
myco2 <- read.csv("CO2.csv", sep = ";") %>% 
    tbl_df() %>%
    select(Blocs,Fermenter,Species,Time = Time..h.,CO2rate = CO2_rate..g.L.h.,CO2 = CO2..g.L.) %>%
    filter(!grepl("Td",Species)) %>%
    filter(Time <= 300)

myco2$Species <- factor(myco2$Species, levels = c( "Ho","Sc vs Ho", "Hu","Sc vs Hu", "Mf", "Sc vs Mf","Mp","Sc vs Mp",  "Sc"))


knitr::kable(head(myco2))

mypopulation <- read.csv("all_data.csv", sep = ";") %>%
    tbl_df() %>%
    filter(Time <= 300) %>%
    select(Blocs,Fermenter,Species,Time, Pop_total,Prop_Sc,viability) %>%
    mutate(PopV = Pop_total*viability) %>%
    mutate(PopV.Sacc = PopV * Prop_Sc, PopV.NonSacc = PopV *(1-Prop_Sc)) %>%
    filter(!grepl("Td",Species)) %>%
    tbl_df
    
mypopulation$Species <- factor(mypopulation$Species, levels = c( "Ho","Sc vs Ho", "Hu","Sc vs Hu", "Mf", "Sc vs Mf","Mp","Sc vs Mp",  "Sc"))





mydata.hplc <- read.csv("all_data.csv", sep = ";") %>%
    tbl_df() %>%
    filter(Time <= 300) %>%
    select(-CO2,-Pop_total,-Prop_Sc,-viability,-Ammonium,-Alanine,-Arginine,-Aspartate,-Glutamine,-Glutamate,-Glycine,-Histidine,-Methionine,-Phenylalanine,-Serine,-Threonine,-Tryptophane,-Tyrosine,-Valine) %>% 
    filter(!grepl("Td",Species)) %>%
    na.omit()

mydata.aa <- read.csv("all_data.csv", sep = ";") %>%
    tbl_df() %>%
    filter(Time <= 300) %>%
    select(-CO2,-Pop_total,-Prop_Sc,-viability) %>%
    select(-glucose,-fructose,-glycerol,-acetate,-ethanol,-alpha,-succinate,-pyruvate) %>%
    filter(!grepl("Td",Species)) %>%
    na.omit()


mydata.aa$Species <- factor(mydata.aa$Species, levels = c( "Ho","Sc vs Ho", "Hu","Sc vs Hu", "Mf", "Sc vs Mf","Mp","Sc vs Mp", "Sc"))

mydata.hplc$Species <- factor(mydata.hplc$Species, levels = c( "Ho","Sc vs Ho", "Hu","Sc vs Hu", "Mf", "Sc vs Mf","Mp","Sc vs Mp", "Sc"))

```






# CO2
We first estimate for each fermentation the Vmax and the MaxCO2

```{r }
stock.para.CO2 <- NULL
for(k in unique(myco2$Blocs)){
    sub.data.1 <- myco2 %>% 
        filter(Blocs == k)
    for(i in unique(sub.data.1$Species)){
        sub.data.2 <- sub.data.1 %>% filter(Species == i)
        for(j in unique(sub.data.2$Fermenter)){
            sub.data <- sub.data.2 %>% filter(Fermenter == j)
            
            Vmax <- max(sub.data$CO2rate)
            tVmax <- sub.data$Time[max(which(sub.data$CO2rate == max(sub.data$CO2rate)))]
            
            maxCO2 <- max(sub.data$CO2)
            latency.co2 <- max(sub.data$Time[sub.data$CO2 <= 1])
            
            stock.para.CO2 <- rbind.data.frame(stock.para.CO2,
                                          data.frame(Bloc = k,
                                                     Species = i,
                                                     Fermenter = j,
                                                     Vmax = Vmax, 
                                                     MaxCO2 = maxCO2))
            
            
            
            
        }
        
    }
}


stock.para.CO2 %>%
    group_by(Species) %>%
    dplyr::summarise(mean.MaxCO2 = mean(MaxCO2),
                     sd.MaxCO2 = sd(MaxCO2),
                     mean.Vmax = mean(Vmax),
                     sd.Vmax = sd(Vmax))

```

## Bloc effect

```{r}
mytab.bloc.test <- stock.para.CO2 %>%
    filter(Species == "Sc") %>%
    select(-Species,-Fermenter)

anova(lm(Vmax~Bloc,mytab.bloc.test))
anova(lm(MaxCO2~Bloc,mytab.bloc.test))

```






Then we can analyse the effect of the species on both parameters and plot the results

## Vmax
```{r}
Vmax.aov <- aov(Vmax~Species,data = stock.para.CO2)
Vmax.tukey <- HSD.test(Vmax.aov,"Species")

anova(lm(MaxCO2~Species,stock.para.CO2))


plot.group(Vmax.tukey, las = 3)
means.Vmax <- Vmax.tukey$means %>%
    select(Vmax,std) %>%
    mutate(ymin = Vmax - std, ymax = Vmax + std)
means.Vmax$Species <- row.names(Vmax.tukey$means)
groups.Vmax <- Vmax.tukey$groups
groups.Vmax$Species <- rownames(groups.Vmax)

means.Vmax <- left_join(means.Vmax,groups.Vmax, by = "Species") %>%
    select(Species,Vmax = Vmax.x, std, ymin, ymax, groups)
    
knitr::kable(means.Vmax, digits = 2)

g.Vmax <- means.Vmax %>%
    ggplot(aes(x = reorder(Species,Vmax), y = Vmax, ymin = ymin, ymax = ymax, label = groups)) +
    geom_bar(stat = "identity") +
    geom_text(col = "red",vjust = 0, hjust = 1) +
    geom_errorbar() +
    theme_tufte() +
    labs(x = "", y = "Vmax (g/L/h)",title = "A)")
```
## MaxCO2
```{r}
MaxCO2.aov <- aov(MaxCO2~Species,data = stock.para.CO2)
MaxCO2.tukey <- HSD.test(MaxCO2.aov,"Species")

anova(lm(MaxCO2~Species,stock.para.CO2))

means.MaxCO2 <- MaxCO2.tukey$means %>%
    select(MaxCO2,std) %>%
    mutate(ymin = MaxCO2 - std, ymax = MaxCO2 + std)
means.MaxCO2$Species <- row.names(MaxCO2.tukey$means)
groups.MaxCO2 <- MaxCO2.tukey$groups
groups.MaxCO2$Species <- rownames(groups.MaxCO2)

means.MaxCO2 <- left_join(means.MaxCO2,groups.MaxCO2, by = "Species") %>%
    select(Species,MaxCO2 = MaxCO2.x, std, ymin, ymax, groups)
    
knitr::kable(means.MaxCO2 %>% arrange(MaxCO2), digits = 2)

g.MaxCO2 <- means.MaxCO2 %>%
    ggplot(aes(x = reorder(Species,MaxCO2), y = MaxCO2, ymin = ymin, ymax = ymax, label = groups)) +
    geom_bar(stat = "identity") +
    geom_text(col = "red",vjust = 0, hjust = 1) +
    geom_errorbar() +
    theme_tufte()+
    labs(x = "", y = "CO2max (g/L)",title = "B)")
```
## Graphics 
```{r}
g.Vmax + g.MaxCO2
```

# Population

```{r }
stock.para.pop <- NULL
par(mfrow = c(3,3))
myerror <- NULL
for(k in unique(mypopulation$Blocs)){
    sub.data.1 <- mypopulation %>% filter(Blocs == k)
    for(i in unique(sub.data.1$Species)){
        sub.data.2 <- sub.data.1 %>% filter(Species == i)
        for(j in unique(sub.data.2$Fermenter)){
            sub.data <- sub.data.2 %>% 
                filter(Fermenter == j) %>%
                select(-Blocs, -Fermenter, -Species)
            
            if(i == "Mf") {
                sub.data$Pop_total[sub.data$Time == 0] <- 1000000
            }
            
            
            mypropfinal <- last(sub.data$Prop_Sc)
            myviability <- last(sub.data$viability)
            
            gc_fit <- SummarizeGrowth(sub.data$Time, sub.data$Pop_total)
            plot(gc_fit)
            
            mymu <- gc_fit$vals$r
            myK = gc_fit$vals$k
            
            stock.para.pop <- rbind.data.frame(stock.para.pop,
                                          data.frame(Bloc = k,
                                                     Species = i,
                                                     Fermenter = j,
                                                     mu = mymu, 
                                                     K = myK, 
                                                     propfinal = mypropfinal,
                                                     viability = myviability))
        }
        
    }
}
```
## Bloc effect

```{r}
mytab.bloc.test <- stock.para.pop %>%
    filter(Species == "Sc") %>%
    select(-Species,-Fermenter,-propfinal)

anova(lm(mu~Bloc,mytab.bloc.test))
anova(lm(log10(K)~Bloc,mytab.bloc.test))
anova(lm(viability~Bloc,mytab.bloc.test))

```




## mu
```{r}
mu.aov <- aov(mu~Species,data = stock.para.pop)
mu.tukey <- HSD.test(mu.aov,"Species")

anova(lm(mu~Species+Bloc,stock.para.pop))

means.mu <- mu.tukey$means %>%
    select(mu,std) %>%
    mutate(ymin = mu - std, ymax = mu + std)
means.mu$Species <- row.names(mu.tukey$means)
groups.mu <- mu.tukey$groups
groups.mu$Species <- rownames(groups.mu)

means.mu <- left_join(means.mu,groups.mu, by = "Species") %>%
    select(Species,mu = mu.x, std, ymin, ymax, groups)

means.mu <- means.mu %>% 
    mutate(mu.red = mu/1000000,std.red = std/1000000)
    
knitr::kable(means.mu, digits = 2)

g.mu <- means.mu %>%
    ggplot(aes(x = reorder(Species,mu), y = mu, ymin = ymin, ymax = ymax, label = groups)) +
    geom_bar(stat = "identity") +
    geom_text(col = "red",vjust = 0, hjust = 1) +
    geom_errorbar() +
    theme_tufte() +
    labs(x = "", y = "mu (cells/mL/h)",title = "A)")
```
## K
```{r}
K.aov <- aov(K~Species,data = stock.para.pop)
K.tukey <- HSD.test(K.aov,"Species")

anova(lm(K~Species+Bloc,stock.para.pop))

means.K <- K.tukey$means %>%
    select(K,std) %>%
    mutate(ymin = K - std, ymax = K + std)
means.K$Species <- row.names(K.tukey$means)
groups.K <- K.tukey$groups
groups.K$Species <- rownames(groups.K)

means.K <- left_join(means.K,groups.K, by = "Species") %>%
    select(Species,K = K.x, std, ymin, ymax, groups)

means.K <- means.K %>% 
    mutate(K.red = K/100000000,std.red = std/100000000)
    
knitr::kable(means.K, digits = 2)

g.K <- means.K %>%
    ggplot(aes(x = reorder(Species,K), y = K, ymin = ymin, ymax = ymax, label = groups)) +
    geom_bar(stat = "identity") +
    geom_text(col = "red",vjust = 0, hjust = 1) +
    geom_errorbar() +
    theme_tufte() +
    labs(x = "", y = "K (cells/mL)",title = "B)")
```
## Prop
```{r}
propfinal.aov <- aov(propfinal~Species,data = stock.para.pop %>% filter(grepl("vs",Species)))
propfinal.tukey <- HSD.test(propfinal.aov,"Species")

anova(lm(propfinal~Species+Bloc,stock.para.pop%>% filter(grepl("vs",Species))))

means.propfinal <- propfinal.tukey$means %>%
    select(propfinal,std) %>%
    mutate(ymin = propfinal - std, ymax = propfinal + std)
means.propfinal$Species <- row.names(propfinal.tukey$means)
groups.propfinal <- propfinal.tukey$groups
groups.propfinal$Species <- rownames(groups.propfinal)

means.propfinal <- left_join(means.propfinal,groups.propfinal, by = "Species") %>%
    select(Species,propfinal = propfinal.x, std, ymin, ymax, groups)


    
knitr::kable(means.propfinal, digits = 2)

g.propfinal <- means.propfinal %>%
    ggplot(aes(x = reorder(Species,propfinal), y = propfinal, ymin = ymin, ymax = ymax, label = groups)) +
    geom_bar(stat = "identity") +
    geom_text(col = "red",vjust = 0, hjust = 1) +
    geom_errorbar() +
    theme_tufte() +
    labs(x = "", y = "Sc final proportion",title = "C)")
```
## Viability
```{r}
viability.aov <- aov(viability~Species,data = stock.para.pop)
viability.tukey <- HSD.test(viability.aov,"Species")

anova(lm(viability~Species+Bloc,stock.para.pop))

means.viability <- viability.tukey$means %>%
    select(viability,std) %>%
    mutate(ymin = viability - std, ymax = viability + std)
means.viability$Species <- row.names(viability.tukey$means)
groups.viability <- viability.tukey$groups
groups.viability$Species <- rownames(groups.viability)

means.viability <- left_join(means.viability,groups.viability, by = "Species") %>%
    select(Species,viability = viability.x, std, ymin, ymax, groups)


    
knitr::kable(means.viability, digits = 2)

g.viability <- means.viability %>%
    ggplot(aes(x = reorder(Species,viability), y = viability, ymin = ymin, ymax = ymax, label = groups)) +
    geom_bar(stat = "identity") +
    geom_text(col = "red",vjust = 0, hjust = 1) +
    geom_errorbar() +
    theme_tufte() +
    labs(x = "", y = "Final viability",title = "D)")
```
## Graphics 

###Parameters
```{r}
g.mu + g.K + g.propfinal + g.viability + plot_layout(ncol = 2)
```

### Dynamcis
```{r}



g.popvivante.ho <- mypopulation %>%
    filter(Blocs == "Sc vs Ho") %>%
    select(Species, Time,PopV) %>%
    group_by(Time,Species) %>% 
    dplyr::summarise(sdt = erreur.std(PopV),
              PopV = mean(PopV)) %>%
    mutate(ymin = PopV - sdt, ymax = PopV + sdt) %>%
    ungroup() %>%
    ggplot(aes(x = Time, y = PopV, color = Species, ymax = ymax, ymin = ymin)) +
    geom_line() +
    geom_linerange() +
    scale_y_continuous(limits=c(1,max(mypopulation$PopV)))+
    scale_x_continuous(limits=c(0,max(mypopulation$Time)))+
    scale_shape_manual(values=c(19,19,20))+
    scale_color_manual(values = c("darkorange1", "darkorange3","black")) +
    scale_linetype_manual( values = c(1,3,1)) +
    theme_tufte()+
    theme(legend.position ="None") +
    labs(x = "Time (h)", y = "Pop (cells/mL)")

g.souspop.ho <- mypopulation %>%
    filter(Species == "Sc vs Ho") %>%
    select(Time,PopV,PopV.Sacc,PopV.NonSacc) %>%
    gather(key = type, val = pop,-Time) %>%
    group_by(Time,type) %>% 
    dplyr::summarise(sdt = erreur.std(pop),
              pop = mean(pop)) %>%
    mutate(ymin = pop - sdt, ymax = pop + sdt) %>%
    ungroup() %>%
    ggplot(aes(x = Time, y = pop, color = type, ymax = ymax, ymin = ymin)) +
    geom_line() +
    geom_linerange() +
    scale_y_continuous(limits=c(1,max(mypopulation$PopV)))+
    scale_x_continuous(limits=c(0,max(mypopulation$Time)))+
    scale_shape_manual(values=c(19,19,20))+
    scale_color_manual(values = c("darkorange3","darkorange1","black")) +
    scale_linetype_manual(values = c(1,2,2)) +
    theme_tufte()+
    theme(legend.position ="None") +
    labs(x = "Time (h)", y = "Pop (cells/mL)")

g.popvivante.hu <- mypopulation %>%
    filter(Blocs == "Sc vs Hu") %>%
    select(Species, Time,PopV) %>%
    group_by(Time,Species) %>% 
    dplyr::summarise(sdt = erreur.std(PopV),
              PopV = mean(PopV)) %>%
    mutate(ymin = PopV - sdt, ymax = PopV + sdt) %>%
    ungroup() %>%
    ggplot(aes(x = Time, y = PopV, color = Species, ymax = ymax, ymin = ymin)) +
    geom_line() +
    geom_linerange() +
    scale_y_continuous(limits=c(1,max(mypopulation$PopV)))+
    scale_x_continuous(limits=c(0,max(mypopulation$Time)))+
    scale_shape_manual(values=c(19,19,20))+
    scale_color_manual(values = c("red", "red3", "black")) +
    scale_linetype_manual( values = c(1,3,1)) +
    theme_tufte()+
    theme(legend.position ="None") +
    labs(x = "Time (h)", y = "Pop (cells/mL)")

g.souspop.hu <- mypopulation %>%
    filter(Species == "Sc vs Hu") %>%
    select(Time,PopV,PopV.Sacc,PopV.NonSacc) %>%
    gather(key = type, val = pop,-Time) %>%
    group_by(Time,type) %>% 
    dplyr::summarise(sdt = erreur.std(pop),
              pop = mean(pop)) %>%
    mutate(ymin = pop - sdt, ymax = pop + sdt) %>%
    ungroup() %>%
    ggplot(aes(x = Time, y = pop, color = type, ymax = ymax, ymin = ymin)) +
    geom_line() +
    geom_linerange() +
    scale_y_continuous(limits=c(1,max(mypopulation$PopV)))+
    scale_x_continuous(limits=c(0,max(mypopulation$Time)))+
    scale_shape_manual(values=c(19,19,20))+
    scale_color_manual(values = c("red3","red","black")) +
    scale_linetype_manual(values = c(1,2,2)) +
    theme_tufte()+
    theme(legend.position ="None") +
    labs(x = "Time (h)", y = "Pop (cells/mL)")

g.popvivante.mp <- mypopulation %>%
    filter(Blocs == "Sc vs Mp") %>%
    select(Species, Time,PopV) %>%
    group_by(Time,Species) %>% 
    dplyr::summarise(sdt = erreur.std(PopV),
              PopV = mean(PopV)) %>%
    mutate(ymin = PopV - sdt, ymax = PopV + sdt) %>%
    ungroup() %>%
    ggplot(aes(x = Time, y = PopV, color = Species, ymax = ymax, ymin = ymin)) +
    geom_line() +
    geom_linerange() +
    scale_y_continuous(limits=c(1,max(mypopulation$PopV)))+
    scale_x_continuous(limits=c(0,max(mypopulation$Time)))+
    scale_shape_manual(values=c(19,19,20))+
    scale_color_manual(values = c("turquoise","turquoise4","black")) +
    scale_linetype_manual( values = c(1,3,1)) +
    theme_tufte()+
    theme(legend.position ="None") +
    labs(x = "Time (h)", y = "Pop (cells/mL)")

g.souspop.mp <- mypopulation %>%
    filter(Species == "Sc vs Mp") %>%
    select(Time,PopV,PopV.Sacc,PopV.NonSacc) %>%
    gather(key = type, val = pop,-Time) %>%
    group_by(Time,type) %>% 
    dplyr::summarise(sdt = erreur.std(pop),
              pop = mean(pop)) %>%
    mutate(ymin = pop - sdt, ymax = pop + sdt) %>%
    ungroup() %>%
    ggplot(aes(x = Time, y = pop, color = type, ymax = ymax, ymin = ymin)) +
    geom_line() +
    geom_linerange() +
    scale_y_continuous(limits=c(1,max(mypopulation$PopV)))+
    scale_x_continuous(limits=c(0,max(mypopulation$Time)))+
    scale_shape_manual(values=c(19,19,20))+
    scale_color_manual(values = c("turquoise4","turquoise","black")) +
    scale_linetype_manual(values = c(1,2,2)) +
    theme_tufte()+
    theme(legend.position ="None") +
    labs(x = "Time (h)", y = "Pop (cells/mL)")


g.popvivante.mf <- mypopulation %>%
    filter(Blocs == "Sc vs Mf") %>%
    select(Species, Time,PopV) %>%
    group_by(Time,Species) %>% 
    dplyr::summarise(sdt = erreur.std(PopV),
              PopV = mean(PopV)) %>%
    mutate(ymin = PopV - sdt, ymax = PopV + sdt) %>%
    ungroup() %>%
    ggplot(aes(x = Time, y = PopV, color = Species, ymax = ymax, ymin = ymin)) +
    geom_line() +
    geom_linerange() +
    scale_y_continuous(limits=c(1,max(mypopulation$PopV)))+
    scale_x_continuous(limits=c(0,max(mypopulation$Time)))+
    scale_shape_manual(values=c(19,19,20))+
    scale_color_manual(values = c("green3","green4","black")) +
    scale_linetype_manual( values = c(1,3,1)) +
    theme_tufte()+
    theme(legend.position ="None") +
    labs(x = "Time (h)", y = "Pop (cells/mL)")

g.souspop.mf <- mypopulation %>%
    filter(Species == "Sc vs Mf") %>%
    select(Time,PopV,PopV.Sacc,PopV.NonSacc) %>%
    gather(key = type, val = pop,-Time) %>%
    group_by(Time,type) %>% 
    dplyr::summarise(sdt = erreur.std(pop),
              pop = mean(pop)) %>%
    mutate(ymin = pop - sdt, ymax = pop + sdt) %>%
    ungroup() %>%
    ggplot(aes(x = Time, y = pop, color = type, ymax = ymax, ymin = ymin)) +
    geom_line() +
    geom_linerange() +
    scale_y_continuous(limits=c(1,max(mypopulation$PopV)))+
    scale_x_continuous(limits=c(0,max(mypopulation$Time)))+
    scale_shape_manual(values=c(19,19,20))+
    scale_color_manual(values = c("green4","green3","black")) +
    scale_linetype_manual(values = c(1,2,2)) +
    theme_tufte()+
    theme(legend.position ="None") +
    labs(x = "Time (h)", y = "Pop (cells/mL)")

g.population <- g.popvivante.ho + g.souspop.ho + g.popvivante.hu + g.souspop.hu + g.popvivante.mp + g.souspop.mp + g.popvivante.mf + g.souspop.mf + plot_layout(ncol = 2)


ggsave(g.population,file="population.png")
```
```{r}
g.population <- g.popvivante.ho + g.souspop.ho + plot_layout(ncol = 1)

read_pptx() %>%
    add_slide(layout = "Title and Content", master = "Office Theme") %>%
    ph_with_vg(code = print(g.population),type = "body") %>%
    print(target="figure population2.pptx")
```

# Resources consumption
## Sugar
```{r}
mydata.sugar <- NULL
for(i in unique(mydata.hplc$Blocs)){
    sub.data.1 <- mydata.hplc %>% filter(Blocs == i)
    for(j in unique(sub.data.1$Fermenter)){
        sub.data <- sub.data.1 %>% 
            filter(Fermenter == j) %>%
            filter(Time == max(Time))
        
        mydata.sugar <- rbind.data.frame(mydata.sugar, data.frame(Blocs = i,Fermenter = j, Species = sub.data$Species,glucose = sub.data$glucose, fructose = sub.data$fructose))
    }
}

mydata.sugar$sugar <- mydata.sugar$glucose + mydata.sugar$fructose

sugar.aov <- aov(sugar~Species,data = mydata.sugar)
sugar.tukey <- HSD.test(sugar.aov,"Species")

anova(lm(sugar~Species+Blocs,mydata.sugar))

means.sugar <- sugar.tukey$means %>%
    select(sugar,std) %>%
    mutate(ymin = sugar - std, ymax = sugar + std)
means.sugar$Species <- row.names(sugar.tukey$means)
groups.sugar <- sugar.tukey$groups
groups.sugar$Species <- rownames(groups.sugar)

means.sugar <- left_join(means.sugar,groups.sugar, by = "Species") %>%
    select(Species,sugar = sugar.x, std, ymin, ymax, groups)
means.sugar$Species <- factor(means.sugar$Species, levels = c( "Ho","Sc vs Ho", "Hu","Sc vs Hu", "Mf", "Sc vs Mf","Mp","Sc vs Mp",  "Sc"))

    
knitr::kable(means.sugar, digits = 2)

g.sugar <- means.sugar %>%
    ggplot(aes(x = Species, y = sugar, ymin = ymin, ymax = ymax, label = groups,fill = Species)) +
    geom_bar(stat = "identity") +
    geom_text(col = "red",vjust = 0, hjust = 1) +
    geom_errorbar() +
    scale_fill_manual(values = c("darkorange","darkorange3","red","red3","green3","green4","turquoise","turquoise3","black")) +
    theme_tufte() +
    labs(x = "", y = "Concentration (g/L)", title = "A)") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), legend.position='none')
```

## Cinétique sucre 

```{r}


mytab <- mydata.hplc %>%
    select(Blocs,Species, Time, glucose, fructose) %>%
    gather(key = sugar, val = val, -Species, -Time, -Blocs) %>%
    group_by(Blocs,Species, Time, sugar) %>%
    dplyr::summarise(mean.val = mean(val,na.rm = T),
                     sd.val = sd(val,na.rm = T)) %>%
    mutate(ymin = mean.val - sd.val, ymax = mean.val + sd.val)



mytab$Species <- factor(mytab$Species, levels = c("Ho","Hu", "Mf", "Mp","Sc vs Ho","Sc vs Hu", "Sc vs Mf",  "Sc vs Mp", "Sc"))


g.cons.sugar <- mytab %>%
    filter(Blocs != "Sc vs Td") %>%
    ggplot(aes(x = Time, y = mean.val, col = Species, ymin = ymin, ymax = ymax, linetype = sugar)) +
    geom_line() + 
    geom_linerange() +
    scale_colour_manual(values = c("darkorange","red","green3","turquoise","darkorange3","red3","green4","turquoise3","black")) +
    facet_wrap(~Species,ncol = 4) +
    theme_tufte() +
    labs(x = "Time (h)", y = "Concentration (g/L)")



g.cons.sugar    
```



## Total NAS

```{r}
mydata.aa$total <- apply(mydata.aa[,c("Ammonium","Alanine","Arginine","Aspartate","Glutamine","Glutamate","Glycine","Histidine","Methionine","Phenylalanine","Serine","Threonine","Tryptophane","Tyrosine","Valine")],1,sum)

stock.cons.aa <- NULL
for(i in unique(mydata.aa$Blocs)){
    sub.data.1 <- mydata.aa %>% filter(Blocs == i)
    for(j in unique(sub.data.1$Fermenter)){
        sub.data.2 <- sub.data.1 %>% filter(Fermenter == j)
        for(k in 5:ncol(sub.data.2)){
            sub.data <- data.frame(sub.data.2$Time,sub.data.2[,k])
            
            delta.aa <- as.numeric(sub.data.2[1,k] - sub.data.2[2,k])
            ratio.aa <- as.numeric(delta.aa/sub.data.2[1,k])
         
            stock.cons.aa <- rbind.data.frame(stock.cons.aa,
                                              data.frame(Blocs = i,
                                                         Species = sub.data.2$Species[1],
                                                         AA = colnames(sub.data.2)[k],
                                                         debut = as.numeric(sub.data.2[1,k]),
                                                         fin = as.numeric(sub.data.2[2,k]),
                                                         delta = delta.aa,
                                                         ratio = ratio.aa))
        }
    }
}


left.NAS.aov <- aov(fin~Species,data = stock.cons.aa %>% filter(AA == "total"))
left.NAS.tukey <- HSD.test(left.NAS.aov,"Species")

anova(lm(fin~Species+Blocs,stock.cons.aa %>% filter(AA == "total")))

means.left.NAS <- left.NAS.tukey$means %>%
    select(fin,std) %>%
    mutate(ymin = fin - std, ymax = fin + std)
means.left.NAS$Species <- row.names(left.NAS.tukey$means)
groups.left.NAS <- left.NAS.tukey$groups
groups.left.NAS$Species <- rownames(groups.left.NAS)

means.left.NAS <- left_join(means.left.NAS,groups.left.NAS, by = "Species") %>%
    select(Species,NAS = fin.x, std, ymin, ymax, groups)
means.left.NAS$Species <- factor(means.left.NAS$Species, levels = c( "Ho","Sc vs Ho", "Hu","Sc vs Hu", "Mf", "Sc vs Mf","Mp","Sc vs Mp",  "Sc"))

    
knitr::kable(means.left.NAS, digits = 2)

g.NAS <- means.left.NAS %>%
    ggplot(aes(x = Species, y = NAS, ymin = ymin, ymax = ymax, label = groups,fill = Species)) +
    geom_bar(stat = "identity") +
    geom_text(col = "red",vjust = 0, hjust = 1) +
    geom_errorbar() +
    scale_fill_manual(values = c("darkorange","darkorange3","red","red3","green3","green4","turquoise","turquoise3","black")) +
    theme_tufte() +
    labs(x = "", y = "Concentration (mg/L)", title = "B)") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), legend.position='none')
```
## Ratio by AA
```{r}

AA.ratio <- stock.cons.aa %>%
    group_by(Species,AA) %>%
    dplyr::summarise(mean = mean(ratio))

myorder.AA <- AA.ratio %>%
    group_by(AA) %>%
    dplyr::summarise(mean = mean(mean)) %>%
    arrange(mean) %>%
    select(AA) %>%
    as.data.frame()
myorder.AA <- as.character(myorder.AA[,1])

myorder.Species <- AA.ratio %>%
    group_by(Species) %>%
    dplyr::summarise(mean = mean(mean)) %>%
    arrange(mean) %>%
    select(Species) %>%
    as.data.frame()
myorder.Species <- as.character(myorder.Species[,1])

AA.ratio$AA <- factor(AA.ratio$AA, levels = myorder.AA)
AA.ratio$Species <- factor(AA.ratio$Species, levels = myorder.Species)

g.aa.ratio <- AA.ratio %>%
    filter(AA != "total") %>%
    ggplot(aes(x = Species,y = AA)) +
    geom_tile(aes(fill = mean)) +
    scale_fill_gradient2(low = "darkgreen",high = "red",na.value = "grey50",midpoint = 0.75) +
    theme_tufte() +
    labs(x="Species",y="Amino Acids",title="C)") +
    geom_text(aes(x=Species,y=AA,label=paste(round(mean*100),"%",sep="")),size = 3) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
```

```{r}
figure3 <- g.sugar + g.NAS - g.aa.ratio + plot_layout(ncol = 1)
figure3
```
```{r}
g.aa.ratio

read_pptx() %>%
    add_slide(layout = "Title and Content", master = "Office Theme") %>%
    ph_with_vg(code = print(g.aa.ratio),type = "body") %>%
    print(target="aa ratio.pptx")


```

#Production total HPLC 
```{r}
stock.prod <- NULL
for(i in unique(mydata.hplc$Blocs)){
    sub.data.1 <- mydata.hplc %>% filter(Blocs == i)
    for(j in unique(sub.data.1$Fermenter)){
        sub.data <- sub.data.1 %>% 
            filter(Fermenter == j) %>%
            filter(Time == max(Time)) %>%
            mutate(sucre.cons = 200 - glucose - fructose) %>%
            select(Blocs, Species, glycerol, acetate, ethanol, alpha, succinate, pyruvate, sucre.cons)
        
       stock.prod <- rbind.data.frame(stock.prod,sub.data) 
    }
}

stock.data <- NULL
for(i in 3:8){
    sub.data <- stock.prod %>%
        select(c(Blocs = 1, Species = 2, metabolite = i))
    
    
    metabolite.aov <- aov(metabolite~Species,data = sub.data)
    metabolite.tukey <- HSD.test(metabolite.aov,"Species")

    print(paste("############ ",colnames(stock.prod)[i]," ############"))
    print(anova(lm(metabolite~Species+Blocs,sub.data)))

    means.metabolite <- metabolite.tukey$means %>%
        select(metabolite,std) %>%
        mutate(ymin = metabolite - std, ymax = metabolite + std)
    means.metabolite$Species <- row.names(metabolite.tukey$means)
    groups.metabolite <- metabolite.tukey$groups
    groups.metabolite$Species <- rownames(groups.metabolite)

    means.metabolite <- left_join(means.metabolite,groups.metabolite, by = "Species") %>%
        select(Species,metabolite = metabolite.x, std, ymin, ymax, groups)
    means.metabolite$Species <- factor(means.metabolite$Species, levels = c( "Ho","Sc vs Ho", "Hu","Sc vs Hu", "Mf", "Sc vs Mf","Mp","Sc vs Mp",  "Sc"))

    
    print(knitr::kable(means.metabolite, digits = 2))

    g.metabolite <- means.metabolite %>%
        ggplot(aes(x = reorder(Species,metabolite),, y = metabolite, ymin = ymin, ymax = ymax, label = groups,fill = Species)) +
        geom_bar(stat = "identity") +
        geom_text(col = "red",vjust = 0, hjust = 1) +
        geom_errorbar() +
        scale_fill_manual(values = c("darkorange","darkorange3","red","red3","green3","green4","turquoise","turquoise3","black")) +
        theme_tufte() +
        labs(x = "", y = "Concentration (g/L)", title = colnames(stock.prod)[i]) +
        theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), legend.position='none')
    print(g.metabolite)
    
    
    
    tab.tmp <- means.metabolite %>%
        mutate(type = colnames(stock.prod)[i]) %>%
        select(Species,type,metabolite,std,groups)
    
    stock.data <- rbind.data.frame(stock.data, tab.tmp)
}


```
```{r}
stock.data$type[stock.data$type == "glycerol"] <- "Glycerol"
stock.data$type[stock.data$type == "succinate"] <- "Succinate"
stock.data$type[stock.data$type == "acetate"] <- "Acetate"
stock.data$type[stock.data$type == "ethanol"] <- "Ethanol"
stock.data$type[stock.data$type == "pyruvate"] <- "Pyruvate"
stock.data$type[stock.data$type == "alpha"] <- "Alpha-ketoglutarate"

g.production <- stock.data %>%
    mutate(ymin = metabolite - std, ymax =  metabolite + std) %>%
    ggplot(aes(x = Species, y = metabolite, ymin = ymin, ymax = ymax, fill = Species, label = groups)) +
    geom_bar(stat = "identity") +
    geom_errorbar() +
    geom_text(col = "black",vjust = 0, hjust = 0.5, aes(x = Species, y = ymax*1.05)) +
    scale_fill_manual(values = c("darkorange","darkorange3","red","red3","green3","green4","turquoise","turquoise3","black")) +
    facet_wrap(~type,scales = "free_y") +
    labs(x = "", y = "Concentration (g/L)") +
    theme_tufte() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5,face = "bold.italic"), legend.position='none',axis.ticks.x=element_blank())
g.production
ggsave(g.production,file="production.png")
```





# Yield relative to S. cerevisiae Yield
```{r}
myyield <- stock.prod %>%
    mutate(Glycerol = glycerol/sucre.cons,
           Acetate = acetate/sucre.cons,
           Ethanol = ethanol/sucre.cons,
           Alpha_ketoglutarate = alpha/sucre.cons,
           Succinate = succinate/sucre.cons,
           Pyruvate = pyruvate/sucre.cons) %>%
    select(Blocs, Species,Glycerol,Acetate,Ethanol,Alpha_ketoglutarate,Succinate,Pyruvate)


tabgather <- myyield  %>%
    gather(key = parametres, val = value,-Species,-Blocs)


mean.Sc <- tabgather %>% 
    filter(Species == "Sc") %>%
    group_by(parametres) %>%
    dplyr::summarise(mean = mean(value))

tabgather$value.r <- 0
for(i in 1:nrow(tabgather)){
    tabgather$value.r[i] <- (tabgather$value[i]-mean.Sc$mean[mean.Sc$parametres == tabgather$parametres[i]])/mean.Sc$mean[mean.Sc$parametres == tabgather$parametres[i]]
    
}

tabgather %>%
    group_by(Species,parametres) %>%
    dplyr::summarise(mean=mean(value.r),sdt = erreur.std(value.r)) %>%
    mutate(ymin = mean - sdt, ymax = mean + sdt) %>%
    ggplot(aes(x = Species, y = mean, ymin = ymin, ymax = ymax, fill = Species)) +
    geom_bar(stat="identity") +
    geom_errorbar() +
    scale_fill_manual(values = c("darkorange","darkorange3","red","red3","green3","green4","turquoise","turquoise3","black")) +
    #facet_wrap(~parametres,scales = "free_y") +
    facet_wrap(~parametres) +
    theme_tufte() +
    theme(axis.text.x = element_text(angle = 90,hjust = 1, vjust = 0.5)) +
    labs(x = "", y = "relative yield to S. cerevisiae yield")

tabgather %>%
    group_by(Species,parametres) %>%
    dplyr::summarise(mean=mean(value.r),sdt = erreur.std(value.r)) %>%
    mutate(ymin = mean - sdt, ymax = mean + sdt) %>%
    ggplot(aes(x = Species, y = mean, ymin = ymin, ymax = ymax, fill = Species)) +
    geom_bar(stat="identity") +
    geom_errorbar() +
    scale_fill_manual(values = c("darkorange","darkorange3","red","red3","green3","green4","turquoise","turquoise3","black")) +
    facet_wrap(~parametres,scales = "free_y") +
    theme_tufte() +
    theme(axis.text.x = element_text(angle = 90,hjust = 1, vjust = 0.5)) +
    labs(x = "", y = "relative yield to S. cerevisiae yield")




```
```{r}
for(i in unique(tabgather$parametres)){
    sub.data <- tabgather %>%
        filter(parametres == i)
    
    
    relative.yield.aov <- aov(value.r~Species,data = sub.data)
    relative.yield.tukey <- HSD.test(relative.yield.aov,"Species")

    print(paste("############ ",colnames(stock.prod)[i]," ############"))
    print(anova(lm(value.r~Species+Blocs,sub.data)))

    means.relative.yield <- relative.yield.tukey$means %>%
        select(value.r,std) %>%
        mutate(ymin = value.r - std, ymax = value.r + std)
    means.relative.yield$Species <- row.names(relative.yield.tukey$means)
    groups.relative.yield <- relative.yield.tukey$groups
    groups.relative.yield$Species <- rownames(groups.relative.yield)

    means.relative.yield <- left_join(means.relative.yield,groups.relative.yield, by = "Species") %>%
        select(Species,value.r = value.r.x, std, ymin, ymax, groups)
    means.relative.yield$Species <- factor(means.relative.yield$Species, levels = c( "Ho","Sc vs Ho", "Hu","Sc vs Hu", "Mf", "Sc vs Mf","Mp","Sc vs Mp",  "Sc"))

    
    print(knitr::kable(means.relative.yield, digits = 2))

    g.relative.yield <- means.relative.yield %>%
        ggplot(aes(x = reorder(Species,value.r), y = value.r, ymin = ymin, ymax = ymax, label = groups,fill = Species)) +
        geom_bar(stat = "identity") +
        geom_text(col = "red",vjust = 0, hjust = 1) +
        geom_errorbar() +
        scale_fill_manual(values = c("darkorange","darkorange3","red","red3","green3","green4","turquoise","turquoise3","black")) +
        theme_tufte() +
        labs(x = "", y = "relative yield to Sc yield", title = i) +
        theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), legend.position='none')
    print(g.relative.yield)
}
```

# ACP 

```{r}
tab.CO2 <- stock.para.CO2 %>%
    group_by(Species) %>%
    dplyr::summarise(Vmax = mean(Vmax), MaxCO2 = mean(MaxCO2))
tab.pop <- stock.para.pop %>%
    group_by(Species) %>%
    dplyr::summarise(mu = mean(mu), K = mean(K), viability = mean(viability))
tab.sugar <- means.sugar %>%
    select(Species,sugar)
tab.NAS <- means.left.NAS %>%
    select(Species, NAS)
tab.prod <- stock.prod %>% 
    group_by(Species) %>%
    dplyr::summarise(glycerol = mean(glycerol),
                     acetate = mean(acetate),
                     ethanol = mean(ethanol), 
                     alpha = mean(alpha),
                     succinate = mean(succinate),
                     pyruvate = mean(pyruvate))



left_join(tab.CO2,tab.pop)
left_join(left_join(tab.CO2,tab.pop),tab.sugar)
left_join(left_join(left_join(tab.CO2,tab.pop),tab.sugar), tab.NAS)
tab.ACP <- left_join(left_join(left_join(left_join(tab.CO2,tab.pop),tab.sugar), tab.NAS),tab.prod) %>%
    as.data.frame()


rownames(tab.ACP) <- tab.ACP$Species
tab.ACP <- tab.ACP[,-1]



library(FactoMineR)

my.ACP <- PCA(tab.ACP, ind.sup = c(3,5,6,9))


```


